Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-16 paper.tex
some comments

Various phenomena in the real world are not evenly distributed across the environment such as watershed, wild fires, traffic, human crowd movement, or wild life movement from insects, fish, birds, and mammals.
To acquire appropriate information about such phenomena, we also need to distribute the sensors accordingly to cover all aspects sufficiently. While manual deployment will mitigate the problem, not all phenomena are known exactly \emph{a priori} and would require maintenance and adjustment during runtime. Even worse, with phenomena able to change their location, size, and distribution---as it is the case with crowds, wild life, or wild fires---an adaptation of the collective sensors is required in order to keep the required information.

@@

In this paper, we propose a novel approach combining aggregate computing~\cite{Beal2015Computer} with Graph Neural Networks~\cite{Zhou2020AIOpen} to focus on the relevant areas. We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
%
This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
Specifically, 
 our approach utilises aggregate computing to disseminate information by manipulating a computational field, 
 which is a distributed data structure that maps devices to information. 
 This field functions as a dynamic environment in which information is continuously updated and diffused throughout the system. 
 By leveraging this layer, we can build collective distributed intelligence using \ac{GNN},
  which uses the information in the field to determine the best action for a given task.
\ga{first draft, need a refinement}
The remainder of this paper is structured as follows. First, we introduce the relevant background and problem formulation in \Cref{sec:background}. Afterwards, we introduce our approach in \Cref{sec:approach}. \Cref{sec:eval} outlines the performed experiments and discusses the obtained results. Finally, we will present our conclusions in \Cref{sec:conclusion}.

@@

@@

@@

 the use of \acp{GNN} allows us to define policies on a variable neighbourhood  that is essential in such systems as this can vary based on the neighbourhood policy,
 and it is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes. 

@@

 During learning, it is as if we have a \emph{single} agent collectively choosing which action to take (i.e., resulting action graph). 

@@

  \While{not done}{
  Observe current graph observations $G_o$;

  With probability $\epsilon$ select a random graph action $G_a$, 
  otherwise select $G_a =$ 
  $\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;

@@

  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G^i_o)$ 
   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$;

@@

\section{Evaluation}
\label{sec:eval}
\ga{Page budget: 2 pages \\}
%\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
%I dunno if it makes sense to discuss the variants (i.e., fixed area, two areas and moving area). We will see.}
To test the effectiveness of the proposed approach, 
 we decided to experiment with a case study related to swarm robotics, 
 specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
 or monitoring the water levels in a canal.

>>>>

Various phenomena in the real world are not evenly distributed across the environment such as watershed, wild fires, traffic, human crowd movement, or wild life movement from insects, fish, birds, and mammals.
To acquire appropriate information about such phenomena, we also need to distribute the sensors accordingly to cover all aspects sufficiently. While manual deployment will mitigate the problem, not all phenomena are known exactly \emph{a priori} and would require maintenance and adjustment during runtime. Even worse, with phenomena able to change their location, size, and shape---as it is the case with crowds, wild life, or wild fires---an adaptation of the collective sensors is required in order to ensure all necessary information is acquired.

@@

In this paper, we propose a novel approach utilising aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}. 
\lukas{can we have a very short summary of how it works here. something along the lines of ``First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent. Second, we utilise a \ac{GNN} within the \ac{DQN}. Here the \ac{GNN} provides node embeddings for each agent, serving as input for the \ac{DQN}. The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.''}
We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
%
This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
Specifically, 
 our approach utilises aggregate computing to disseminate information by manipulating a computational field, 
 which is a distributed data structure that maps devices to information. 
 This field functions as a dynamic environment in which information is continuously updated and diffused throughout the system. 
 By leveraging this layer, we can build collective distributed intelligence using \ac{GNN},
  which uses the information in the field to determine the best action for a given task.
\ga{first draft, need a refinement}
The remainder of this paper is structured as follows. First, we introduce the relevant background and problem formulation in \Cref{sec:background}. Afterwards, we introduce our approach in \Cref{sec:approach}. \Cref{sec:eval} outlines the performed experiments and discusses the obtained results. Finally, we will present our conclusions in \Cref{sec:conclusion}.

@@

\lukas{maybe: ``While the graph, containing the aggregate information, might appear as global knowledge, this is not the case as the information is never aggregated globally. The individual nodes only combine information from their local neighbourhood.''}

@@

 \lukas{Global information is not available to the agent, right? The agent only provide locally aggregated information (resulting from the AC part)?}

@@

 the use of \acp{GNN} allows us to define policies on a variable neighbourhood, which is essential in such systems as this can change due to the applied neighbourhood policy.
 It is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes~\cite{??}. 

@@

 During learning, it is as if we have a \emph{single} agent collectively choosing which action to take (i.e., resulting action graph). \lukas{this makes it sound like a singe agent has all information, selecting the actions for all other agents...}

@@

  \While{not done}{
  Observe current graph observations $G_o$; \lukas{whats $G_o$?}

%  $r \leftarrow $ randomNumber();
%  
%	\If{$r \leq \epsilon$}{select a random graph action $G_a$}
%	\Else{select $G_a =$\\
%		$\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;}
%  
  With probability $\epsilon$ select a random graph action $G_a$, 
  otherwise select $G_a =$ 
  $\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;

@@

  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G^i_o)$ 
   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$; \lukas{how does the merge work? do we explain this?}

@@

\section{Evaluation}
\label{sec:eval}

%% Put three subfigures of simulations, start, after, end
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Initial state}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Simulation of the proposed approach. 
  The dots represent the robots, 
  the circle area on the top right represent the phenomenon to be monitored, 
  and the gray area represent the area covered by the robots.}
  \label{fig:simulations}
\end{figure}
\ga{Page budget: 2 pages \\}
%\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
%I dunno if it makes sense to discuss the variants (i.e., fixed area, two areas and moving area). We will see.}
To test the effectiveness of the proposed approach, 
 we decided to experiment with a case study related to swarm robotics, 
 specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
 or monitoring the water levels in a canal. The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
 \lukas{``Initially we consider a stationary phenomenon before expanding towards a moving phenomenon. Additionally, the phenomenon can change its shape during runtime. Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function. Figure~\ref{fig:simulations} illustrates a set of 25 agents and a single, stationary phenomenon with a with a 3D Gaussian distribution.''}
